---
title: "Sleep Disorder"
output: powerpoint_presentation
author: "Sarah Alshumayri, Reema Abdallah, Yehya Aseeri" 
date: "2023-11-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) 
```

## Problem Statement

"Exploring the Interplay Between Sleep, Health, and Lifestyle: A Data-Driven Approach"

This project investigates the relationship between sleep patterns (efficiency, duration, stages) and various health and lifestyle factors (physical activity, stress, caffeine/alcohol consumption, sleep disorders), aiming to uncover correlations and insights into how these elements impact an individual's sleep quality and overall well-being.

## Background

Sleep is a critical component of overall health and well-being, yet it is often influenced by a myriad of factors ranging from lifestyle choices to health conditions. Understanding the dynamics of sleep and its relationship with different health and lifestyle factors can provide crucial insights for improving sleep quality and, by extension, overall health and life satisfaction.

This project leverages data-driven analysis to explore these relationships. With datasets detailing sleep efficiency, stages of sleep, and other relevant sleep metrics, combined with data on physical activity, stress levels, BMI categories, and more, we can draw a comprehensive picture of the multifaceted nature of sleep in the context of contemporary lifestyles. This analysis can potentially lead to actionable recommendations for individuals seeking to improve their sleep and overall health.

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)

# Read the dataset
sleep_data <- read.csv("Data/Sleep_health_and_lifestyle_dataset.csv")

# Preliminary data exploration
summary(sleep_data)

# Basic visualization
ggplot(sleep_data, aes(x = Age, y = Sleep.Duration)) +
  geom_point() +
  labs(title = "Age vs. Sleep Duration")
```

## Methods Explored

Reema's part

## The Methods/Tools Used
Yahya's part


## Results 
Sarah's Part

### Further Results:
Sarah's Part

## Lessons Learned
Reema's part

## Future Plans
Reema's part

